A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2018-11-25 01_Introduction.Rmd
started on some words about research history

The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017]. Archaeology has a special perspective to offer here: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018]. Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous remains of human live. Similarity between artefacts within and beyond cultural-historical contexts are traditionally explained by interaction: We rightfully expect humans to learn production as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them. The relatedness of historically and spatially close material remains generally is a result of long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].

>>>>

The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017]. This theoretical framework is as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* to *Neoevolutionism* up until the ninteen sixties and seventies [@petermannGeschichteEthnologie2004].

Archaeology has a special perspective to offer here: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018]. Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous remains of human live. Similarity between artefacts within and beyond cultural-historical contexts are traditionally explained by interaction: We rightfully expect humans to learn production as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them. The relatedness of historically and spatially close material remains generally is a result of long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
